% paper/sections/04_experimental_setup.tex
%
% First draft. Structure follows paper/04_experimental_setup.md (four
% subsections: inventory → agents → seeds & inference unit → metrics,
% cost rates, exclusions). Methodology lives in §3; §4 reports only
% the scope of what was actually run. Numbers resolve through
% \result / \respct / \resp macros backed by paper/data/*.csv; literal
% versions/sample sizes (claude-sonnet-4-6, gpt-5.5, n=93, n=100, 3 seeds)
% are whitelist-allowed in paper/lint.py.
%
% Forward references: \S\ref{sec:method-arms} (§3.1 three-arm tool surface),
% \S\ref{sec:method-harness} (§3.2 harness design / subprocess isolation /
% wall-time cap), \S\ref{sec:m-integrity} (§3.3 SWE-bench integrity),
% \S\ref{sec:m-cost} (§3.5 cost methodology), \S\ref{sec:m-artifact}
% (§3.6 artifact-suite contract).

% ──────────────────────── §4.1 Benchmark inventory ─────────────────────
\subsection{Benchmark inventory}
\label{sec:setup-inventory}

We evaluate on two benchmarks chosen to cover disjoint task regimes; the
methodology for each lives in \S\ref{sec:m-integrity} (SWE-bench protocol)
and \S\ref{sec:m-artifact} (artifact-suite contract), and is not
repeated here.

\paragraph{Artifact suite (computation regime).} A self-contained suite of
$n{=}\result{artifact_categories}{total:n}$ tasks across nine categories,
three at $n{=}\result{artifact_categories}{data_engineering:n}$ (data
engineering, data science, ML engineering) and six at
$n{=}\result{artifact_categories}{algorithmic:n}$ (algorithmic, data
processing, enumeration, iterative numerical, stateful reasoning,
verification-heavy). Each task is a
\texttt{(workspace, hidden grader, reference output)} triple under the
contract of \S\ref{sec:m-artifact}; the reference-output round-trip
self-test passes on all $\result{artifact_categories}{total:n}$ tasks in
CI.

\paragraph{SWE-bench Mini (modification regime).} A curated
$n{=}\result{swebench_repos}{total:total:n}$ subset of
SWE-bench~\cite{jimenez2024swebench} in two sub-corpora.
\emph{verified-mini}
($n{=}\result{swebench_repos}{verified-mini:total:n}$): Django
($\result{swebench_repos}{verified-mini:django:n}$), Sphinx
($\result{swebench_repos}{verified-mini:sphinx:n}$), drawn from
SWE-bench Verified~\cite{report:openai2024swebenchverified}.
\emph{datasci-mini}
($n{=}\result{swebench_repos}{datasci-mini:total:n}$): sklearn
($\result{swebench_repos}{datasci-mini:scikit-learn:n}$), matplotlib
($\result{swebench_repos}{datasci-mini:matplotlib:n}$), xarray
($\result{swebench_repos}{datasci-mini:xarray:n}$), sympy
($\result{swebench_repos}{datasci-mini:sympy:n}$), seaborn
($\result{swebench_repos}{datasci-mini:seaborn:n}$), astropy
($\result{swebench_repos}{datasci-mini:astropy:n}$). The split spans the
canonical modification regime and the data-science / scientific-Python
tail underrepresented in SWE-bench Verified.

% ──────────────────────────── §4.2 Agents ──────────────────────────────
\subsection{Agents}
\label{sec:setup-surfaces}

We instantiate the three-arm ablation of \S\ref{sec:method-arms}
(\texttt{baseline}, \texttt{bash\_only}, \texttt{code\_only}) on two
production coding agents:
\emph{Claude Code}~\cite{repo:anthropic2026claudecode}
(model \texttt{claude-\allowbreak sonnet-\allowbreak 4-6}, Claude Code 2.1.139) and
\emph{Codex CLI}~\cite{repo:openai2026codex} (model \texttt{gpt-5.5}).
Both are invoked under the per-arm subprocess-isolation protocol of
\S\ref{sec:method-harness} (per-run isolated config dir, no
session-persistence, sandbox/permission bypass). Codex is reported as
a co-headline finding, not a footnote-level generalisation probe: the
agent-design contrast on SWE-bench
(Codex $p=$\,\resp{fig.02_signflip}{swebench-codex:p} vs.\ Claude
$p=$\,\resp{fig.02_signflip}{swebench-claude:p} on the
\texttt{code\_only} cost contrast) is one of the four cells that
Table~\ref{tab:code-only-headline} in \S\ref{sec:results} reports.

\paragraph{Codex apply\_patch leakage audit.} The
\S\ref{sec:method-arms} soft-disable was audited on all $110$ Codex
SWE-bench JSONLs (seed~1): \texttt{file\_change} events fire $7$ times
on \texttt{bash\_only} and $4$ times on \texttt{code\_only} (both
${<}1\%$, the $4$ concentrated in
\texttt{scikit-learn\allowbreak\_\_scikit-learn-11596}) vs.\ $344$ on
\texttt{baseline}. The cross-agent comparisons in \S\ref{sec:results}
are not biased by leaked \texttt{apply\_patch} use.
% TODO macro: \result{setup}{codex_cli_version} — blocked on
% paper/data/setup_constants.csv promotion. Pin verbatim once landed.

% ──────────────────── §4.3 Seeds and inference unit ────────────────────
\subsection{Seeds and the unit of inference}
\label{sec:setup-inference}

\paragraph{Seeds.} Each \texttt{(instance, arm)} cell is run with
3 seeds, drawn as independent harness invocations on disjoint
container snapshots. Per-seed raw records and the analysis scripts
that produced every \verb|\result{}| macro in this paper are released with
the harness; we do not reproduce them in this PDF.

\paragraph{The unit of inference is the task, not the seed.} A
seed-marginal standard error (SE) would pool over instances of dramatically heterogeneous
difficulty and confuse between-task variance with within-task
replication noise. Empirically, on Claude SWE-bench the cross-instance
SD of cache-adjusted cost is roughly an order of magnitude larger than
the paired per-task contrast SE, and pairing flips several headline
cells from NS to significant.

\paragraph{Procedure.} For every comparison reported in
\S\ref{sec:results}: (i) collapse the 3 seeds within each
\texttt{(task, arm)} cell to a per-task mean; (ii) for each ordered arm
pair $(A, B)$ within a benchmark, compute the per-task vector of
differences $\Delta_t = \mathrm{mean}_A(t) - \mathrm{mean}_B(t)$ and
report mean $\Delta$, $\mathrm{SE}_\Delta = \mathrm{SD}(\Delta) /
\sqrt{n_{\text{tasks}}}$, and a 95\% normal-approximation CI
($n_{\text{tasks}} \in \{93, 100\}$ supports the approximation); (iii)
pair the CI with a paired Wilcoxon signed-rank
$p$~\cite{wilcoxon1945} on the $\Delta$ vector for continuous metrics
(cost, input tokens, output tokens). Wilcoxon is preferred over the
paired $t$-test because the per-task $\Delta$ distributions are
heavy-tailed---a handful of pathological SWE-bench instances dominate
parametric variance while the median direction is clear in every cell.
Pass rate uses the same machinery on per-task pass rates
$\in \{0, 1/3, 2/3, 1\}$, a Wilcoxon-on-rates analogue of McNemar that
respects the within-task seed structure. The marginal per-arm summary
(the ``row'' in Table~\ref{tab:code-only-headline}) is the mean over
tasks of the per-task mean, $\pm \mathrm{SD}_{\text{task}} /
\sqrt{n_{\text{tasks}}}$---the SE that respects task as the unit of
replication. Per-seed values are derivable from the released data but
are not the inferential unit. Implementation:
\texttt{paper/\allowbreak data/\allowbreak scripts/\allowbreak paired\allowbreak\_contrasts.py}, fully reproducible
from the released JSONL records.

% ──────────────────── §4.4 Metrics, cost rates, exclusions ─────────────
\subsection{Metrics, cost rates, and exclusions}
\label{sec:setup-metrics}

\paragraph{Metric surface.} Table~\ref{tab:code-only-headline} in
\S\ref{sec:results} reports four columns per \texttt{(benchmark, agent)}
cell. \emph{Pass rate}, defined as $\mathrm{PASS} / (\mathrm{PASS} +
\mathrm{FAIL})$ with \texttt{env\_fail} instances excluded from the
denominator (see below), reported as an absolute $\Delta$ in
percentage points. \emph{Cache-adjusted cost}: the
median-floor-adjusted variant defined in \S\ref{sec:m-cost}, which
charges the shared per-arm prefix at the cache-read rate using the
per-arm median first-turn \texttt{cache\allowbreak\_read\allowbreak\_input\allowbreak\_tokens} as the
floor; reported as a relative $\Delta$\%. \emph{Input tokens} and
\emph{output tokens}: per-run sums of the corresponding usage fields
across user-role API turns (cached and uncached pooled for input);
reported as relative $\Delta$\%. Turns, dollars per turn, and median
per-instance cost are diagnostic rather than headline; the raw token-based
cost is the underlying input to the cache-adjusted column, not a
parallel reporting axis (\S\ref{sec:m-cost}).

\paragraph{Cost rates.} The cache-adjusted cost formula plugs in the
vendor-published per-million-token rates for input, cached-input, and
output, separately for \texttt{claude-\allowbreak sonnet-\allowbreak 4-6} and \texttt{gpt-5.5},
as of the access date in \S\ref{sec:m-cost}. Pinning the rate constants
in the setup section is what makes every cost column in
\S\ref{sec:results} reproducible from the released JSONL records.
% TODO macro: per-Mtok rate table via \result{cost_rates}{...}
% — blocked on paper/data/cost_rates.csv promotion.

\paragraph{Wall budget.} Each $(\text{instance}, \text{arm}, \text{run})$
is capped at 60 minutes of wall time (\S\ref{sec:method-harness});
per-task execution-budget fields reserved in the artifact task schema
are not enforced.

