% paper/sections/99_limitations.tex
% Required by the ACM template (desk-reject if missing). Roughly 0.25 page
% (about five bullets); 0.5 ceiling per outline.md §7.
%
% Per the 2026-05-28 outline decision, the analyze/ pathology pipeline is
% NOT listed as a limitation — it is future work, not a caveat on the
% claims this paper actually makes.

\begin{itemize}

\item \textbf{Two agent surfaces.} The crossed
$3\,{\times}\,2\,{\times}\,2$ design exercises Claude Code
(\texttt{claude-sonnet-4-6}, v2.1.139) and Codex CLI (\texttt{gpt-5.5})
and no other agents. GPT-5, Gemini~2.5, and open-weight coding scaffolds
(Qwen-Coder, DeepSeek-Coder, and similar) are out of scope. The
three-mechanism account in \S\ref{sec:discussion} predicts where each
component should generalise---batching is API-shape-dependent,
edit-friction scales with edit volume on any agent that lacks a native
byte-precise editor, and failure-cost scales with the unsolvable-instance
fraction---but the \emph{joint} four-cell footprint is verified only on
the two agents reported here.

\item \textbf{Sample size and power.} Per-cell $n$ is $93$ for Artifact
and $100$ for SWE-bench Mini, three independent seeds each. Two of the
four headline \texttt{code\_only}-vs-rival contrasts are significant at
$p < 10^{-8}$; the SWE-bench/Claude cost gap of
$\respct{paired_contrasts}{swebench:claude:onlycode-vs-baseline:cost_adj:mean_delta}{swebench:claude:onlycode-vs-baseline:cost_adj:mean_b}$
($p = \resp{paired_contrasts}{swebench:claude:onlycode-vs-baseline:cost_adj:wilcoxon_p}$)
is reported as directional rather than significant, and a
SWE-bench-Verified-scale replication is reserved for follow-up work.

\item \textbf{Artifact-suite scope.} Tasks are single-language (Python)
on Linux, with category coverage skewed toward analytical, numerical,
and ML-flavoured computation. The grader contract in
\S\ref{sec:method-artifact} is language-agnostic but has not been
exercised on Rust, systems-engineering tasks, or long-horizon
multi-day scenarios, so the four-cell pattern's stability outside
that scope is an open empirical question.

\item \textbf{Cost is cache-adjusted, not cache-isolated.} Neither the
Anthropic Messages API nor the OpenAI Responses API exposes a
documented knob that disables, scopes, or partitions the server-side
prompt cache or any related token-caching primitive
(\S\ref{sec:method-cost}). We therefore charge every token at the
non-cached rate and floor each instance's first-turn
\texttt{cache\_read} at the per-arm median to normalise the
system-prompt warm/cold lottery without claiming isolation beyond it.
Within-task multi-turn cache reads and any deeper server-side state we
cannot observe are not adjusted; the residual gap to a true
steady-state production cost is therefore unmeasured.

\item \textbf{Single-shot, internal-surface, post-\#287 protocol.} Each
$(\text{instance}, \text{arm}, \text{seed})$ triple is a single agent
invocation: multi-shot prompting, self-consistency, and
scaffold-evolution variants are out of scope, as are external MCP tool
surfaces (Drive, Slack, vendor APIs) of the kind studied by the
industry write-ups we position against (\S\ref{sec:related}). On the
modification benchmark we report only the post-Issue~\#287 protocol
with post-agent \texttt{test\_patch} application
(\S\ref{sec:method-integrity}); pre-\#287 legacy data is not
comparable and is omitted entirely. Twelve SWE-bench instances
(\texttt{sympy} + \texttt{mwaskom}) returned API auth-failures on
seed~1 and were re-run on the same agent version.

\end{itemize}
